Predictions for detecting BHNS and other double compact objects binaries with LISA

### 2021-03-20 paper/tex/paper.tex
overhaul references, add previous studies table

\usepackage{lipsum}
\usepackage{physics}
\usepackage{multirow}
\usepackage{xspace} %this package handles when to use a whitespace in math symbols
\usepackage{natbib} %this package can be used for creating an automated bibliography.

@@

%Latex allows you to define abbreviations. \newcommand{.. here comes the short notation ...}{ ... here comes the full latex command... small change}

@@

% Define symbols for Rsun and Msun, Lsun and Teff ensuring proper formatting
\newcommand{\Rsun}{\ensuremath{\,\mathrm{R_\odot}}\xspace} % Solar radius symbol
\newcommand{\Msun}{\ensuremath{\,\mathrm{M_\odot}}\xspace} % Solar mass symbol
\newcommand{\Lsun}{\ensuremath{\,\mathrm{M_\odot}}\xspace} % Solar luminosity symbol
\newcommand{\Teff}{\ensuremath{\,\mathrm{T_{\rm eff}}}\xspace} % Effective temperature symbol
\newcommand{\Myr}{\ensuremath{\,\mathrm{Myr}}\xspace} % symbol for Mega year
\newcommand{\Gyr}{\ensuremath{\,\mathrm{Gyr}}\xspace} % symbol for Gyr

@@

\begin{abstract}
{The Laser Inteferometer Space Antenna (LISA) presents an exciting new lens through which to view the realm of gravitational waves. Detections of double compact objects with LISA may help to constrain uncertainties in binary evolution and understand the prevalence of electromagnetic counterparts to gravitational wave events. We estimate the number of Galactic DCOs that will be detected by LISA using a sample generated with the rapid population synthesis code COMPAS and a model of the Milky Way that includes not only the radial birth profiles but also the star formation and enrichment histories. We explore the effects of varying 15 physical assumptions that focus on key uncertainties in our understanding of mass transfer, common envelope and supernova physics. For our fiducial model, we find that over a 4 year mission, LISA will detect 26 BHBHs, 27 BHNSs and 12 NSNSs. These detection rates increase to 41, 45 and 19 respectively for a 10 year mission length. We present the distributions of the characteristics of the detectable binaries and explore how our different models affect them. We also discuss how well we can distinguish between different DCO types and methods for doing so.}
\end{abstract}

@@

Since the first direct detection of gravitational waves by the LIGO scientific collaboration in 2015 \citep{LIGO_2016_BBHinfirstRun}, there have been 50 gravitational wave events detected by LIGO \citep{LIGO_GWTC1, LIGO_GWTC2}. The investigation of double compact object (DCO) population statistics provides an essential tool for constraining uncertainties in binary evolution and predicting distributions of observable parameters for different DCO types.

@@

One particularly interesting and elusive gravitational wave source is a black hole neutron star binary (BHNS). Of all the events detected by LIGO, none can be confidently attributed to the merger of a black hole and a neutron star, though several events such as GW190425 and GW190814 have not been ruled out as a BHNS merger \citep{LIGO_GW190425, LIGO_GW190814}.

@@

The detection of double compact objects with LISA has been investigated in many recent studies. \citet{Korol+2017, Korol+2020} use the binary population synthesis code SeBa to focus on predictions for double white dwarf binaries (WDWD). \citet{Lamberts+2018, Lamberts+2019} instead use a hydrodynamic simulation from the FIRE project \citep{Hopkins+2014} to produce predictions for double black hole binaries (BHBH) and WDWD. Closer to this study, \citet{Breivik+2019} uses the COSMIC population synthesis code to predict detection rates for all double compact objects and \citet{Lau+2020} uses the same population synthesis code as this study, COMPAS, and focusses on double neutron star binaries (NSNS).

@@

Each binary is sampled from initial distributions for the primary and secondary masses as well as the separation. The primary mass, or the mass of the initially more massive star, is restricted to $m_1 \in [5, 150] \unit{M_{\odot}}$ and drawn from the \citet{Kroupa2001} Initial Mass Function (IMF), therefore $p(m_1) \propto m_1^{-2.3}$. The secondary mass is drawn using the mass ratio of the binary, which we assume to be uniform on $[0, 1]$, therefore $p(q) = 1$ \citep[consistent e.g. with][]{Sana+2012}. We additionally restrict $m_2 \ge 0.1 \unit{M_{\odot}}$, since this is approximately the minimal mass for a main sequence star. We assume that the initial separation follows a flat in the log distribution with $p(a_i) \propto 1 / a_i$ and $a_i \in [0.01, 1000] \unit{AU}$ \citep{Opik1924, Abt1983}. We assume that all binaries are circular at birth to reduce the dimensions of initial parameters. Since we focus on post-interaction binaries which will have circularised during mass transfer this is a reasonable assumption and is likely not critical for predicting detection rates \citep{Hurley+2002, de-Mink+2015}.

@@

\textit{Common Envelope:} A common envelope phase follows dynamically unstable mass transfer and we parameterise this using the $\alpha$-$\lambda$ prescription from \citet{Webbink1984} and \citet{de-Kool1990}. We assume $\alpha = 1$, such that all of the gravitational binding energy is available for the ejection of the envelope. For $\lambda$ we use the fitting formulae from \citet{Xu+2010, Xu+2010a}. We assume that any Hertzsprung gap donor stars that initiate a common envelope phase will not survive this phase due to a lack of a steep density gradient between the core and envelope \citep{Taam&Sandquist2000, Ivanova+2004}. This follows the `pessimistic' common envelope scenario \citep[c.f.][]{Belczynski+2007a}. We remove any binaries where the secondary immediately fills its Roche lobe upon the conclusion of the common envelope phase as we treat these as failed common envelope ejections.

@@

We assume that stars with helium core masses between $1.6$--$2.25 \unit{M_{\odot}}$ \citep{Hurley+2002} experience electron-capture supernovae \citep{Nomoto1984, Nomoto1987, Ivanova+2008}. We set all remnant masses to $1.26 \unit{M_{\odot}}$ in this case as an approximation of the solution to Equation 8 of \citet{Timmes+1996}. For these supernovae, we set $\sigma_{\rm rms}^{\rm 1D} = 30 \unit{km}{s^{-1}}$ as in \citet{Pfahl+2002} and \citet{Podsiadlowski+2004}.

@@

We assume that kicks are isotropic in the frame of the collapsing star. The maximum neutron star mass is predicted to be between $2$-$3 \unit{M_{\odot}}$ \citep[e.g.][]{Kalogera+1996_maxNS, Fryer+2015, Margalit&Metzger2017}. % removed Lawrence+2015 Alsing+2018
We adopt a maximum neutron star mass of $2.5 \unit{M_{\odot}}$ for the fiducial model and change the \citet{Fryer+2012} prescription accordingly.

@@

\begin{table}[htb]
    \centering
    \begin{tabular}{cl}
        \hline \hline
        Model & Physics Variation \\
        \hline \hline
        \modFid & -- \\
        \hline
        \modBetaLow & Fixed mass transfer efficiency of $\beta=0.25$ \\ 
        \modBetaMed & Fixed mass transfer efficiency of $\beta=0.5$  \\ 
        \modBetaHigh & Fixed mass transfer efficiency of $\beta=0.75$ \\ 
        \modCaseBB & Case BB mass transfer is always unstable \\
        \hline
        \modAlphaLow & CE efficiency parameter $\alpha = 0.5$ \\
        \modAlphaHigh & CE efficiency parameter $\alpha = 2$   \\
        \modOpt & HG donor stars initiating a CE survive CE \\
        \hline
        \modRapid & Fryer rapid SN remnant mass prescription \\
        \modNSLow & Maximum NS mass is fixed to $2\Msun$ \\
        \modNSHigh & Maximum NS mass is fixed to $3\Msun$ \\
        \modNoPISN & PISN and pulsational-PISN not implemented \\
        \modSigLow & $\sigma_{\rm{rms}}^{\rm{1D}}=100 \unit{km}{s^{-1}}$ for core-collapse supernova \\  
        \modSigLower & $\sigma_{\rm{rms}}^{\rm{1D}}=30  \unit{km}{s^{-1}}$ for core-collapse supernova \\ 
        \modNoBH & Black holes receive no natal kick \\
        \hline \hline
    \end{tabular}%
    \caption{Adapted from \citet[][Table 2]{Broekgaarden+2021}, a description of the 15 binary population synthesis models used in this study. \modFid{} is the fiducial model, \modRangeMT{} change mass transfer physics, \modRangeCE{} change common envelope physics and \modRangeSN{} change supernova physics.}
    \label{tab:physics_variations}
\end{table}

@@

For the Milky Way thin disc, we use the model from \citet{Frankel+2018}. The authors developed this model in order to measure the global efficiency of radial migration in the Milky Way. They did this by fitting the model to a sample of red clump stars measured with APOGEE \citep{APOGEE} and Gaia \citep{Gaia-Collaboration+2016}. The result of this work is a model for the Milky Way that is calibrated to observations and includes the star formation history, radial birth profile and chemical enrichment history of the galaxy. We can therefore use this model to place binaries randomly in space and time and match these positions to a metallicity.

The star formation history is given by the following  \citep[][Eq. 2]{Frankel+2018}
\begin{equation}\label{eq:galaxy_tau}
    p(\tau) \propto \exp \qty(-\frac{(\tau_m - \tau)}{\tau_{\rm SFR}}),
\end{equation}
where $\tau$ is the lookback time (the amount of time elapsed between the binary's zero-age main sequence and today), $\tau_m = 12 \unit{Gyr}$ is the assumed age of the Milky Way and $\tau_{\rm SFR} = 6.8 \unit{Gyr}$ was fit by \citet{Frankel+2018}. The radial birth profile of the galaxy is conditional upon its age and is given as \citep[][Eq. 5]{Frankel+2018}
\begin{equation}\label{eq:galaxy_R}
    p(R_0 \vert \tau) = \exp(-\frac{R_0}{R_{\rm exp}(\tau)}) \cdot \frac{R_0}{[R_{\rm exp}(\tau)]^2},
\end{equation}
where $R_0$ is the birth Galactocentric radius, and\footnote{In $R_{\rm exp}(\tau)$, we use 4 kpc instead of 3 kpc for the 0 Gyr exponential scale-length of the disc as it provides a better fit (correspondence with Neige Frankel)}
\begin{equation}
    R_{\rm exp}(\tau) = 4 \unit{kpc} \qty(1 - \alpha_{R_{\rm exp}} \qty(\frac{\tau}{8 \unit{Gyr}})),
\end{equation}
where $\alpha_{R_{\rm exp}} = 0.3$ was fit by \citet{Frankel+2018}. Finally, the metallicity-radius-time relation is given as \citep[][Eq. 7]{Frankel+2018}
\begin{equation}\label{eq:galaxy_FeH}
    \begin{split}
        [{\rm Fe} / {\rm H}] (R, \tau) = &F_m + \nabla [{\rm Fe} / {\rm H}] R \\
        &- \qty(F_m + \nabla [{\rm Fe} / {\rm H}] R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} ) f(\tau),
    \end{split}
\end{equation}
where
\begin{equation}
    f(\tau) = \qty(1 - \frac{\tau}{\tau_m})^{\gamma_{[{\rm Fe} / {\rm H}]}},
\end{equation}
$F_m = -1 \unit{dex}$ is the metallicity of the gas at the center of the disk at $\tau = \tau_m$ and $\nabla [{\rm Fe} / {\rm H}] = -0.075 \unit{kpc^{-1}}$, $R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} = 8.7 \unit{kpc}$ and $\gamma_{[{\rm Fe} / {\rm H}]} = 0.3$ are fit by \citet{Frankel+2018}. We can convert this to our representation of metallicity using \citep[e.g][]{Bertelli+1994}
\begin{equation}\label{eq:galaxy_FeH_to_Z}
    \log Z = 0.977 [{\rm Fe} / {\rm H}] + \log(Z_\odot).
\end{equation}
Additionally, since the Frankel model does not treat the vertical structure of the Milky Way disc, we can supplement it with the vertical distribution from \citet{McMillan2011},
\begin{equation}\label{eq:galaxy_z}
    p(\abs{z}) = \frac{1}{z_d} \exp\qty(-\frac{z}{z_d}),
\end{equation}
where $z$ is the height above the Galactic plane and $z_d = 0.3 \unit{kpc}$. Equations \ref{eq:galaxy_tau}-\ref{eq:galaxy_z} summarise the model that we use for the Milky Way disc and we illustrate these distributions in Fig.~\ref{MAKEME}.

@@

For each Milky Way instance, we randomly sample the following set of parameters
\begin{equation}
    \mathbf{g}_{{i}} = \{\tau, R, Z, z, \theta\}
\end{equation}
for $i = 1, 2, \dots, N_{\rm MW}$, where we set $N_{\rm MW} = 10^{5}$, $\tau, R, Z$ and $z$ are defined and sampled using Eq.\,\ref{eq:galaxy_tau}-\ref{eq:galaxy_z}, $\theta$ is the polar angle sampled uniformly on $[0, 2\pi)$ and $Z$ is the metallicity. Figure~\ref{fig:galaxy_metallicity} shows an example of a random Milky Way instance created with these distributions. This shows how these distributions translate to positions in the Milky and illustrates the gradient in metallicity over radius.

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{galaxy_diagram.png}
    \caption{A schematic showing how we create a mock Mily Way disc. On the left, we show the various distributions presented in \citet{Frankel+2018} and \citet{McMillan2011}. On the right, we show an example instance of the Milky Way with $10^5$ binaries shown as points. The colour of each point represents its metallicity (with $Z \in [0.0001, 0.022]$). The top panel shows a side-on view and the bottom panel shows a view from above.}
    \label{fig:galaxy_schematic}
\end{figure*}

@@

Each binary is likely to move from its birth orbit. Although all stars in the Galactic disc experience some amount of radial migration \citep{Sellwood&Binney2002, Frankel+2018}, double compact objects generally experience stronger dynamical evolution as a result of the effects of both Blaauw kicks from instantaneous symmetric mass loss \citep{Blaauw1961} and natal kicks from asymmetric supernovae \citep{Hobbs+2005}.

@@

Each binary loses orbital energy to gravitational waves throughout its lifetime. This causes the binary to shrink and circularise over time. In order to assess the detectability of a binary, we need to know its eccentricity and frequency at the time of the LISA mission. For each binary in our simulated Milky Way, we know that the time from DCO formation to today is $\tau - t_{\rm evolve}$ and that the initial eccentricity and semi-major axis are $e_{\rm DCO}$ and $a_{\rm DCO}$. We find the eccentricity of the binary at the start of the LISA mission, $e_{\rm LISA}$, by numerically integrating its time derivative \citep[][Eq. 5.13]{Peters1964} given the initial conditions. This additionally can be converted to the semi-major axis at the start of LISA, $a_{\rm LISA} $\citep[][Eq. 5.11]{Peters1964}, which in turn gives the orbital frequency, $f_{\rm orb, LISA}$, by Kepler's third law.

@@

We define a binary as detectable if its gravitational wave signal has a signal-to-noise ratio of greater than 7 \citep[e.g.][]{Breivik+2019, Korol+2020}. The sky-, polarisation- and orientation-averaged signal-to-noise ratio, $\rho$, of an inspiraling binary can be calculated with the following\footnote{For full derivations of the equations in this section, please see Appendix \ref{app:strain_snr_deriv}.} \citep[e.g.][]{Finn&Thorne2000}
\begin{equation}\label{eq:snr}
    \rho^2 = \sum_{n=1}^{\infty} \int_{f_{n, i}}^{f_{n, f}} \frac{h_{c, n}^{2}}{f_{n}^{2} S_{\rm n}\left(f_{n}\right)} \dd{f_n},
\end{equation}
where $n$ is a harmonic of the gravitational wave signal, $f_n = n \cdot f_{\rm orb}$ is the frequency of the $n^{\rm th}$ harmonic of the gravitational wave signal, $f_{\rm orb}$ is the orbital frequency, $S_{\rm n}(f_n)$ is the LISA sensitivity curve at frequency $f_n$ \citep[e.g.][]{Robson+2019} and $h_{c,n}$ is the characteristic strain of the $n^{\rm th}$ harmonic, given by
\begin{equation}\label{eq:charstrain}
    h^2_{c,n} = \frac{2^{5/3}}{3 \pi^{4/3}} \frac{(G \mathcal{M}_c)^{5/3}}{c^3 D_L^2} \frac{1}{n f_{\rm orb}^{1/3}} \frac{g(n,e)}{F(e)}
\end{equation}
\citep[e.g.][]{Barack&Cutler2004}, where $D_L$ is the luminosity distance to the source, $f_{\rm orb}$ is the orbital frequency, $g(n, e)$ and $F(e)$ are given in \citet{Peters1963} and $\mathcal{M}_c$ is the chirp mass, defined as
\begin{equation}\label{eq:chirp_mass}
    \mathcal{M}_c = \frac{(m_1 m_2)^{3/5}}{(m_1 + m_2)^{1/5}}.
\end{equation}

@@

Firstly, eccentricity enhances the rate of energy emission via gravitational waves by a factor
\begin{equation}
    F(e) = \frac{1 + (73 / 24) e^2 + (37 / 96) e^4}{(1 - e^2)^{7/2}},
\end{equation}
compared to a circular binary with the same semi-major axis \citep[][Eq.\,17]{Peters1963}. This means that eccentric binaries will not only inspiral faster, but also produce gravitational waves with stronger strain amplitudes (when summed over all harmonics).

Secondly, eccentric binaries will emit gravitational waves at many harmonic frequencies (unlike circular binaries, which only emit at twice the orbital frequency). This leads to the gravitational wave signal being diluted over many frequencies higher than the orbital frequency, where the higher the eccentricity, the more harmonics are required to capture all of the gravitational luminosity \citep[see][Fig.\,3]{Peters1963}.

@@

\begin{figure*}[p]
    \centering
    \includegraphics[width=\textwidth]{dco_detections.pdf}
    \caption{The number of expected detections in the LISA mission for different DCO types and model variations. Error bars show the 1 (solid) and 2 (dotted) $\sigma$ uncertainties. The left axis and grid lines show the number of detections in a four year LISA mission and the right axis shows an approximation of the number of detections in a 10 year mission (we scale the axis by $\sqrt{T_{\rm obs}}$, see Table~\ref{tab:detection_rates} for exact rates). Each model is described in further detail in Table~\ref{tab:physics_variations} and details of the fiducial assumptions are in  Section~\ref{sec:fiducial_physics}}
    \label{fig:model_variation_detections}
\end{figure*}

@@

\software{COMPAS (version 02.12.00) \url{http://github.com/TeamCOMPAS/COMPAS}. \citep{Stevenson+2017, Vigna-Gomez+2018, Broekgaarden+2019}, \texttt{Python} available from \url{python.org}, matplotlib \citep{matplotlib},  \texttt{NumPy} \citep{numpy}, 
Astropy \citep[\url{http://www.astropy.org}][]{Astropy-Collaboration+2013,Astropy-Collaboration+2018}.}

@@

>>>>

\usepackage{lipsum}
\usepackage{physics}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{natbib}

@@

@@

@@

\begin{abstract}
{The Laser Interferometer Space Antenna (LISA) presents an exciting new lens through which to view the realm of gravitational waves. Detections of double compact objects with LISA may help to constrain uncertainties in binary evolution and understand the prevalence of electromagnetic counterparts to gravitational wave events. We estimate the number of Galactic DCOs that will be detected by LISA using a sample generated with the rapid population synthesis code COMPAS and a model of the Milky Way that includes not only the radial birth profiles but also the star formation and enrichment histories. We explore the effects of varying 15 physical assumptions that focus on key uncertainties in our understanding of mass transfer, common envelope and supernova physics. For our fiducial model, we find that over a 4 year mission, LISA will detect 26 BHBHs, 27 BHNSs and 12 NSNSs. These detection rates increase to 41, 45 and 19 respectively for a 10 year mission length. We present the distributions of the characteristics of the detectable binaries and explore how our different models affect them. We also discuss how well we can distinguish between different DCO types and methods for doing so.}
\end{abstract}

@@

Since the first direct detection of gravitational waves by the LIGO scientific collaboration in 2015 \citep{Abbott+2016_first_detection}, there have been 50 gravitational wave events detected by LIGO \citep{Abbott+2019_GWTC1,Abbott+2020_GWTC2}. The investigation of double compact object (DCO) population statistics provides an essential tool for constraining uncertainties in binary evolution and predicting distributions of observable parameters for different DCO types.

@@

One particularly interesting and elusive gravitational wave source is a black hole neutron star binary (BHNS). Of all the events detected by LIGO, none can be confidently attributed to the merger of a black hole and a neutron star, though several events such as GW190425 and GW190814 have not been ruled out as a BHNS merger \citep{Abbott+2020_GW190425,Abbott+2020_GW190814}.

@@

The detection of double compact objects with LISA has been investigated in many previous studies through a combination of population synthesis and Milky Way modelling. Several focus only on the detection of the more numerous population of double white dwarf population

The detection of double compact objects with LISA has been investigated in many recent studies. \citet{Korol+2017, Korol+2020} use the binary population synthesis code SeBa to focus on predictions for double white dwarf binaries (WDWD). \citet{Lamberts+2018, Lamberts+2019} instead use a hydrodynamic simulation from the FIRE project \citep{Hopkins+2014} to produce predictions for double black hole binaries (BHBH) and WDWD. Closer to this study, \citet{Breivik+2020} uses the COSMIC population synthesis code to predict detection rates for all double compact objects and \citet{Lau+2020} uses the same population synthesis code as this study, COMPAS, and focusses on double neutron star binaries (NSNS).

@@

Each binary is sampled from initial distributions for the primary and secondary masses as well as the separation. The primary mass, or the mass of the initially more massive star, is restricted to $m_1 \in [5, 150] \unit{M_{\odot}}$ and drawn from the \citet{Kroupa+2001} Initial Mass Function (IMF), therefore $p(m_1) \propto m_1^{-2.3}$. The secondary mass is drawn using the mass ratio of the binary, which we assume to be uniform on $[0, 1]$, therefore $p(q) = 1$ \citep[consistent e.g. with][]{Sana+2012}. We additionally restrict $m_2 \ge 0.1 \unit{M_{\odot}}$, since this is approximately the minimal mass for a main sequence star. We assume that the initial separation follows a flat in the log distribution with $p(a_i) \propto 1 / a_i$ and $a_i \in [0.01, 1000] \unit{AU}$ \citep{Opik+1924, Abt+1983}. We assume that all binaries are circular at birth to reduce the dimensions of initial parameters. Since we focus on post-interaction binaries which will have circularised during mass transfer this is a reasonable assumption and is likely not critical for predicting detection rates \citep{Hurley+2002, deMink+2015}.

@@

\textit{Common Envelope:} A common envelope phase follows dynamically unstable mass transfer and we parameterise this using the $\alpha$-$\lambda$ prescription from \citet{Webbink+1984} and \citet{deKool+1990}. We assume $\alpha = 1$, such that all of the gravitational binding energy is available for the ejection of the envelope. For $\lambda$ we use the fitting formulae from \citet{Xu+2010, Xu+2010a}. We assume that any Hertzsprung gap donor stars that initiate a common envelope phase will not survive this phase due to a lack of a steep density gradient between the core and envelope \citep{Taam+2000, Ivanova+2004}. This follows the `pessimistic' common envelope scenario \citep[c.f.][]{Belczynski+2007}. We remove any binaries where the secondary immediately fills its Roche lobe upon the conclusion of the common envelope phase as we treat these as failed common envelope ejections.

@@

We assume that stars with helium core masses between $1.6$--$2.25 \unit{M_{\odot}}$ \citep{Hurley+2002} experience electron-capture supernovae \citep{Nomoto+1984, Nomoto+1987, Ivanova+2008}. We set all remnant masses to $1.26 \unit{M_{\odot}}$ in this case as an approximation of the solution to Equation 8 of \citet{Timmes+1996}. For these supernovae, we set $\sigma_{\rm rms}^{\rm 1D} = 30 \unit{km}{s^{-1}}$ as in \citet{Pfahl+2002} and \citet{Podsiadlowski+2004}.

@@

We assume that kicks are isotropic in the frame of the collapsing star. The maximum neutron star mass is predicted to be between $2$-$3 \unit{M_{\odot}}$ \citep[e.g.][]{Kalogera+1996, Fryer+2015, Margalit+2017}.
We adopt a maximum neutron star mass of $2.5 \unit{M_{\odot}}$ for the fiducial model and change the \citet{Fryer+2012} prescription accordingly.

@@

\begin{table}[htb]
    \centering
    \begin{tabular}{cl}
        \hline \hline
        Model & Physics Variation \\
        \hline \hline
        \modFid & -- \\
        \hline
        \modBetaLow & Fixed mass transfer efficiency of $\beta=0.25$ \\ 
        \modBetaMed & Fixed mass transfer efficiency of $\beta=0.5$  \\ 
        \modBetaHigh & Fixed mass transfer efficiency of $\beta=0.75$ \\ 
        \modCaseBB & Case BB mass transfer is always unstable \\
        \hline
        \modAlphaLow & CE efficiency parameter $\alpha = 0.5$ \\
        \modAlphaHigh & CE efficiency parameter $\alpha = 2$   \\
        \modOpt & HG donor stars initiating a CE survive CE \\
        \hline
        \modRapid & Fryer rapid SN remnant mass prescription \\
        \modNSLow & Maximum NS mass is fixed to $2\unit{M_{\rm odot}}$ \\
        \modNSHigh & Maximum NS mass is fixed to $3\unit{M_{\rm odot}}$ \\
        \modNoPISN & PISN and pulsational-PISN not implemented \\
        \modSigLow & $\sigma_{\rm{rms}}^{\rm{1D}}=100 \unit{km}{s^{-1}}$ for core-collapse supernova \\  
        \modSigLower & $\sigma_{\rm{rms}}^{\rm{1D}}=30  \unit{km}{s^{-1}}$ for core-collapse supernova \\ 
        \modNoBH & Black holes receive no natal kick \\
        \hline \hline
    \end{tabular}%
    \caption{Adapted from \citet[][Table 2]{Broekgaarden+2021}, a description of the 15 binary population synthesis models used in this study. \modFid{} is the fiducial model, \modRangeMT{} change mass transfer physics, \modRangeCE{} change common envelope physics and \modRangeSN{} change supernova physics.}
    \label{tab:physics_variations}
\end{table}

@@

For the Milky Way thin disc, we use the model from \citet{Frankel+2018}. The authors developed this model in order to measure the global efficiency of radial migration in the Milky Way. They did this by fitting the model to a sample of red clump stars measured with APOGEE \citep{Majewski+2017} and Gaia \citep{GaiaCollaboration+2016}. The result of this work is a model for the Milky Way that is calibrated to observations and includes the star formation history, radial birth profile and chemical enrichment history of the galaxy. We can therefore use this model to place binaries randomly in space and time and match these positions to a metallicity.

The star formation history is given by the following  \citep[][Eq. 2]{Frankel+2018}
\begin{equation}\label{eq:galaxy_tau}
    p(\tau) \propto \exp \qty(-\frac{(\tau_m - \tau)}{\tau_{\rm SFR}}),
\end{equation}
where $\tau$ is the lookback time (the amount of time elapsed between the binary's zero-age main sequence and today), $\tau_m = 12 \unit{Gyr}$ is the assumed age of the Milky Way and $\tau_{\rm SFR} = 6.8 \unit{Gyr}$ was fit by \citet{Frankel+2018}. The radial birth profile of the galaxy is conditional upon its age and is given as \citep[][Eq. 5]{Frankel+2018}
\begin{equation}\label{eq:galaxy_R}
    p(R_0 \vert \tau) = \exp(-\frac{R_0}{R_{\rm exp}(\tau)}) \cdot \frac{R_0}{[R_{\rm exp}(\tau)]^2},
\end{equation}
where $R_0$ is the birth Galactocentric radius, and\footnote{In $R_{\rm exp}(\tau)$, we use 4 kpc instead of 3 kpc for the 0 Gyr exponential scale-length of the disc as it provides a better fit (correspondence with Neige Frankel)}
\begin{equation}
    R_{\rm exp}(\tau) = 4 \unit{kpc} \qty(1 - \alpha_{R_{\rm exp}} \qty(\frac{\tau}{8 \unit{Gyr}})),
\end{equation}
where $\alpha_{R_{\rm exp}} = 0.3$ was fit by \citet{Frankel+2018}. Finally, the metallicity-radius-time relation is given as \citep[][Eq. 7]{Frankel+2018}
\begin{equation}\label{eq:galaxy_FeH}
    \begin{split}
        [{\rm Fe} / {\rm H}] (R, \tau) = &F_m + \nabla [{\rm Fe} / {\rm H}] R \\
        &- \qty(F_m + \nabla [{\rm Fe} / {\rm H}] R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} ) f(\tau),
    \end{split}
\end{equation}
where
\begin{equation}
    f(\tau) = \qty(1 - \frac{\tau}{\tau_m})^{\gamma_{[{\rm Fe} / {\rm H}]}},
\end{equation}
$F_m = -1 \unit{dex}$ is the metallicity of the gas at the center of the disk at $\tau = \tau_m$ and $\nabla [{\rm Fe} / {\rm H}] = -0.075 \unit{kpc^{-1}}$, $R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} = 8.7 \unit{kpc}$ and $\gamma_{[{\rm Fe} / {\rm H}]} = 0.3$ are fit by \citet{Frankel+2018}. We can convert this to our representation of metallicity using \citep[e.g][]{Bertelli+1994}
\begin{equation}\label{eq:galaxy_FeH_to_Z}
    \log Z = 0.977 [{\rm Fe} / {\rm H}] + \log(Z_\odot).
\end{equation}
Additionally, since the Frankel model does not treat the vertical structure of the Milky Way disc, we can supplement it with the vertical distribution from \citet{McMillan+2011},
\begin{equation}\label{eq:galaxy_z}
    p(\abs{z}) = \frac{1}{z_d} \exp\qty(-\frac{z}{z_d}),
\end{equation}
where $z$ is the height above the Galactic plane and $z_d = 0.3 \unit{kpc}$. Equations \ref{eq:galaxy_tau}-\ref{eq:galaxy_z} summarise the model that we use for the Milky Way disc and we illustrate these distributions in Fig.~\ref{fig:galaxy_schematic}.

@@

For each Milky Way instance, we randomly sample the following set of parameters
\begin{equation}
    \mathbf{g}_{{i}} = \{\tau, R, Z, z, \theta\}
\end{equation}
for $i = 1, 2, \dots, N_{\rm MW}$, where we set $N_{\rm MW} = 10^{5}$, $\tau, R, Z$ and $z$ are defined and sampled using Eq.\,\ref{eq:galaxy_tau}-\ref{eq:galaxy_z}, $\theta$ is the polar angle sampled uniformly on $[0, 2\pi)$ and $Z$ is the metallicity. Figure~\ref{fig:galaxy_schematic} shows an example of a random Milky Way instance created with these distributions. This shows how these distributions translate to positions in the Milky and illustrates the gradient in metallicity over radius.

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{galaxy_diagram.png}
    \caption{A schematic showing how we create a mock Milky Way disc. On the left, we show the various distributions presented in \citet{Frankel+2018} and \citet{McMillan+2011}. On the right, we show an example instance of the Milky Way with $10^5$ binaries shown as points. The colour of each point represents its metallicity (with $Z \in [0.0001, 0.022]$). The top panel shows a side-on view and the bottom panel shows a view from above.}
    \label{fig:galaxy_schematic}
\end{figure*}

@@

Each binary is likely to move from its birth orbit. Although all stars in the Galactic disc experience some amount of radial migration \citep{Sellwood+2002, Frankel+2018}, double compact objects generally experience stronger dynamical evolution as a result of the effects of both Blaauw kicks from instantaneous symmetric mass loss \citep{Blaauw+1961} and natal kicks from asymmetric supernovae \citep{Hobbs+2005}.

@@

Each binary loses orbital energy to gravitational waves throughout its lifetime. This causes the binary to shrink and circularise over time. In order to assess the detectability of a binary, we need to know its eccentricity and frequency at the time of the LISA mission. For each binary in our simulated Milky Way, we know that the time from DCO formation to today is $\tau - t_{\rm evolve}$ and that the initial eccentricity and semi-major axis are $e_{\rm DCO}$ and $a_{\rm DCO}$. We find the eccentricity of the binary at the start of the LISA mission, $e_{\rm LISA}$, by numerically integrating its time derivative \citep[][Eq. 5.13]{Peters+1964} given the initial conditions. This additionally can be converted to the semi-major axis at the start of LISA, $a_{\rm LISA} $\citep[][Eq. 5.11]{Peters+1964}, which in turn gives the orbital frequency, $f_{\rm orb, LISA}$, by Kepler's third law.

@@

We define a binary as detectable if its gravitational wave signal has a signal-to-noise ratio of greater than 7 \citep[e.g.][]{Breivik+2020, Korol+2020}. The sky-, polarisation- and orientation-averaged signal-to-noise ratio, $\rho$, of an inspiraling binary can be calculated with the following\footnote{For full derivations of the equations in this section, please see Appendix \ref{app:strain_snr_deriv}.} \citep[e.g.][]{Finn+2000}
\begin{equation}\label{eq:snr}
    \rho^2 = \sum_{n=1}^{\infty} \int_{f_{n, i}}^{f_{n, f}} \frac{h_{c, n}^{2}}{f_{n}^{2} S_{\rm n}\left(f_{n}\right)} \dd{f_n},
\end{equation}
where $n$ is a harmonic of the gravitational wave signal, $f_n = n \cdot f_{\rm orb}$ is the frequency of the $n^{\rm th}$ harmonic of the gravitational wave signal, $f_{\rm orb}$ is the orbital frequency, $S_{\rm n}(f_n)$ is the LISA sensitivity curve at frequency $f_n$ \citep[e.g.][]{Robson+2019} and $h_{c,n}$ is the characteristic strain of the $n^{\rm th}$ harmonic, given by
\begin{equation}\label{eq:charstrain}
    h^2_{c,n} = \frac{2^{5/3}}{3 \pi^{4/3}} \frac{(G \mathcal{M}_c)^{5/3}}{c^3 D_L^2} \frac{1}{n f_{\rm orb}^{1/3}} \frac{g(n,e)}{F(e)}
\end{equation}
\citep[e.g.][]{Barack+2004}, where $D_L$ is the luminosity distance to the source, $f_{\rm orb}$ is the orbital frequency, $g(n, e)$ and $F(e)$ are given in \citet{Peters+1963} and $\mathcal{M}_c$ is the chirp mass, defined as
\begin{equation}\label{eq:chirp_mass}
    \mathcal{M}_c = \frac{(m_1 m_2)^{3/5}}{(m_1 + m_2)^{1/5}}.
\end{equation}

@@

Firstly, eccentricity enhances the rate of energy emission via gravitational waves by a factor
\begin{equation}
    F(e) = \frac{1 + (73 / 24) e^2 + (37 / 96) e^4}{(1 - e^2)^{7/2}},
\end{equation}
compared to a circular binary with the same semi-major axis \citep[][Eq.\,17]{Peters+1963}. This means that eccentric binaries will not only inspiral faster, but also produce gravitational waves with stronger strain amplitudes (when summed over all harmonics).

Secondly, eccentric binaries will emit gravitational waves at many harmonic frequencies (unlike circular binaries, which only emit at twice the orbital frequency). This leads to the gravitational wave signal being diluted over many frequencies higher than the orbital frequency, where the higher the eccentricity, the more harmonics are required to capture all of the gravitational luminosity \citep[see][Fig.\,3]{Peters+1963}.

@@

@@

\software{COMPAS (version 02.12.00) \url{http://github.com/TeamCOMPAS/COMPAS}. \citep{Stevenson+2017, Vigna-Gomez+2018, Broekgaarden+2019}, \texttt{Python} available from \url{python.org}, matplotlib \citep{Hunter+2007},  \texttt{NumPy} \citep{vanderWalt+2011}, 
Astropy \citep[\url{http://www.astropy.org}][]{AstropyCollaboration+2013, AstropyCollaboration+2018}.}

@@

\section{Extra figures}
